### 2016-08-30 Chapters/ContributionsSummary.tex
Switch to B5 paper
Update Makefile

The growing complexity of quantum chemical program packages requires that an
appropriate strategy be devised to implement new features.
Scalability is of paramount importance, but it has become clear that
maintainability and extensibility of the code play an equally important role in
managing this growing complexity.~\autocite{Hatton1994-us, Hatton1997-ml,
Hatton1997-bb, Ioannidis2005-aj, Merali2010-uy, Prinz2011-kz, Wilson2014-vh}

The use of a \emph{modular programming paradigm} has been recognized as
beneficial in many other scientific computing contexts. New features are
isolated into libraries that can be accessed by host programs through a
well-defined \ac{API}.
Computational tasks are thus implemented into separate, independent and
interchangeable modules. These are \emph{developed} and \emph{tested}
independently of the particular host program in which they will be
used.~\cite{Dijkstra1968-zp, Parnas1972-im}
A well-defined \ac{API} clearly delimits the boundaries of the functionality
offered, thus forcing a programming style and standardisation of the
functionality, eventually.~\autocite{Reddy2011-sd}

When coupled with \emph{open-source} licensing, modularity is a proper
step towards ensuring reproducibility of results from scientific
simulation software \cite{Gezelter2015-gz, Krylov2015-fs, Jacob2016-oq}.
Open-source development can fully leverage the benefits
of widespread, cloud-based, free, code development services, such as
hosted \acp{DVCS}, \ac{CI}, code
coverage, static and dynamic code analyses, nightly regression testing,
public issue tracking, code review and so forth:
adoption of a modern code development workflow is easily within reach.

@@

@@

\begin{figure}[!htb]
\centering
\scalebox{0.8}{\input{gfx/algorithm.tikz}}
\caption[Modular approach to programming a \acs{PCM} functionality into an existing \acs{SCF} code.]{
  Schematic view of the implemented SCF algorithm. Computations/data in
  \textcolor{Blue}{blue} are on the \pcmsolver side, in
  \textcolor{Green}{green} on the \DIRAC side.
  }
\label{fig:algorithm}
\end{figure}

The description of chemical phenomena in complex environments, such as
solutions or proteins, is as challenging as it is important
for an accurate prediction of a wide range of properties.~\autocite{Reichardt2010-le}
  Quantum chemistry is nowadays a valuable
tool: its simulations can provide significant insight into complex
chemical phenomena.~\autocite{Nobel1998, Nobel2013}.
The major challenge lies in providing a faithful, realistic, yet cost-effective
description of the environment of the molecule:
a molecule in solution can be represented by a quantum mechanical model
comprising a great number of atoms, thousands or more. Such a simulation would,
however, be impossible to carry out with the current state-of-the-art
computational resources.
Models must be devised to overcome the dimensionality ``disease'' and
the idea of introducing different approximation levels for different
parts of the system under study is particularly appealing, as it helps
our chemical understanding. This is the idea at the basis of
\emph{multiscale} models.

\begin{figure}[!htb]
  \centering
  \includegraphics[width=.45\textwidth]{multiscale.jpg}
\end{figure}

\emph{Continuum} (or \emph{implicit}) models, deal explicitly
only with the degrees of freedom associated with the solute, while
replacing the solvent with a structureless continuum characterized by
its bulk properties \cite{Onsager1936-wf}. Among these models, the \acl{PCM} (\acs{PCM})
is the most widespread and successful \cite{Tomasi2005-vm, Mennucci2008-xl}.

The \acs{PCM} is also an ideal candidate for the creation of a solvation \acs{API}.
Its input from and output to \emph{any} quantum chemistry code is
limited and well-defined.
This provides a natural \acs{API} design: the \acs{API} functions can be compared
\emph{vis-à-vis} with the working equations derived for the different
quantum chemical methods.

>>>>

The growing complexity of quantum chemical program packages requires that an
appropriate strategy be devised to implement new features.
Scalability is of paramount importance, but it has become clear that
maintainability and extensibility of the code play an equally important role in
managing software complexity\autocite{Wilson2014-vh} and ensuring
scientific reproducibility.~\autocite{Hatton1994-us, Hatton1997-ml,
Hatton1997-bb, Ioannidis2005-aj, Merali2010-uy, Prinz2011-kz}

The idea of a \emph{modular programming paradigm} can be traced back to
the dawn of computer science. \citeauthor{Dijkstra1968-zp}
successfully used it in the development of the THE operating
system~\autocite{Dijkstra1968-zp} already in 1968, while
\citeauthor{Parnas1972-im} advocated it as a best practice already in
1972.~\autocite{Parnas1972-im}
Modularity is nothing more than a \emph{divide et impera} strategy
applied to programming and it has been recognized as beneficial in many
other scientific computing contexts.
New features are isolated into libraries that can be accessed by host
programs through a well-defined \ac{API}.
Computational tasks are thus implemented into separate, independent and
interchangeable modules. These are \emph{developed}, \emph{tested},
\emph{packaged} and \emph{distributed}
independently of the particular host program in which they will be
used.
A well-defined \ac{API} clearly delimits the boundaries of the functionality
offered, thus forcing a programming style and standardisation of the
functionality, eventually.~\autocite{Reddy2011-sd}

When coupled with \emph{open-source} licensing, modularity is a proper
step towards ensuring reproducibility of results from scientific
simulation software.~\autocite{Gezelter2015-gz, Krylov2015-fs,
Jacob2016-oq}
Open-source development can fully leverage the benefits of widespread,
cloud-based, free, code development services, such as hosted \acp{DVCS},
continuous integration (CI), code coverage, static and dynamic code
analyses, nightly regression testing, public issue tracking, code review
and so forth: adoption of a modern code development workflow is easily
within reach.

@@

The \acs{PCM} is an ideal candidate for the creation of a solvation
\acs{API}.
Consider for example the schematic representation of a
\acs{PCM}-\acs{SCF} algorithm given in Figure
Its input from and output to \emph{any} quantum chemistry code is
limited and well-defined.
This provides a natural \acs{API} design: the \acs{API} functions can be compared
\emph{vis-à-vis} with the working equations derived for the different
quantum chemical methods.

\begin{figure}[!htb]
\centering
\scalebox{0.8}{\input{gfx/algorithm.tikz}}
\caption[Modular approach to programming a \acs{PCM} functionality into an existing \acs{SCF} code.]{
  Schematic view of the implemented SCF algorithm. Computations/data in
  \textcolor{PMS2229}{blue} are on the \pcmsolver side, in
  \textcolor{Green}{green} on the \DIRAC side.
  }
\label{fig:algorithm}
\end{figure}

@@
